---
title: "Property Sale Predictor"
author: "Bharath Sepuri"
date: "November 2, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

##replacing missing values and null values of Sales data with NA
##filtering the most recent sales data
##replacing missing values and null values of PropertyOwnerLoan data with NA
##merging sales and propertyownerloan on PropertyID


```{r 1}

Sales <- read.csv(file = "sales_final.csv", stringsAsFactors =  FALSE)
Sales[Sales == ""] <- NA
Sales[Sales == "NULL"] <- NA
str(Sales)


install.packages("dplyr")
library(dplyr)
Sales_filtered <- Sales %>%
                    group_by(PropertyID) %>%  
                       arrange(desc(TransactionDate)) %>%
                         do(head(.,n=1))
#saveRDS(Sales_filtered, file = "Sales_filtered_full_clean.rds")


PropertyOwnerLoan <-read.csv(file = "pol_final.csv",stringsAsFactors = FALSE)
PropertyOwnerLoan[PropertyOwnerLoan == ""] <- NA
PropertyOwnerLoan[PropertyOwnerLoan == "NULL"] <- NA
#saveRDS(PropertyOwnerLoan, file = "PropertyOwnerLoan_full_clean.rds")


#sales <- readRDS("Sales_filtered_full_clean.rds")
#propertyownerloan <- readRDS("PropertyOwnerLoan_full_clean.rds")

PropertySales <-merge(sales, propertyownerloan , by = "PropertyID")
PropertySales<-PropertySales[PropertySales$PropertyType=="Multifamily",]

#saveRDS(PropertySales, file = "PropertySales_full_clean.rds")

```

##selecting the most potential predictors and saving the final data
##manipulating the variables to maintain data consistency

```{r 2}

#PropertySales<-readRDS("PropertySales_full_clean.rds")

PropertySales_final<-PropertySales[,c("PropertyID","TransactionType","NOIAmount","MaturityDate","TransactionDate", "LastTransactionDate","LastTransactionAmount","NumUnits","OrigLoanAmt","LoanCount","WAInterestRate","LoanTermMonths","TotSqrFt","RentableSqrFt","AverageRentAmt", "CurOccupancyPct")]

#saveRDS(PropertySales_final, file = "PropertySales_final.rds")

PropertySales_final<-readRDS("PropertySales_final.rds")
PropertySales_final <- PropertySales_final[-which(PropertySales_final$LoanTermMonths < 0),]
PropertySales_final <- PropertySales_final[-which(PropertySales_final$TransactionType =="Entity Level"),]

```

##converting strings as integers
##converting strings as factors

```{r 3}

PropertySales_final[, c(
  "NOIAmount",
  "NumUnits",
  "OrigLoanAmt",
  "LastTransactionAmount",
  "LoanCount",
  "WAInterestRate",
  "LoanTermMonths",
  "TotSqrFt",
  "RentableSqrFt",
  "CurOccupancyPct",
  "AverageRentAmt"
)] <-
  lapply(PropertySales_final[, c(
    "NOIAmount",
    "NumUnits",
    "OrigLoanAmt",
    "LastTransactionAmount",
    "LoanCount",
     "WAInterestRate",
    "LoanTermMonths",
    "TotSqrFt",
    "RentableSqrFt",
    "CurOccupancyPct",
    "AverageRentAmt"
  )], as.integer)


PropertySales_final$TransactionType <- factor(PropertySales_final$TransactionType)
str(PropertySales_final)

```

##changing the data type of dates to date format
##replacing missing values of MaturityDate with LastTransactiondate+10

```{r 4}

PropertySales_final$TransactionDate <- as.Date(PropertySales_final$TransactionDate,format = "%m/%d/%Y")
PropertySales_final$LastTransactionDate <- as.Date(PropertySales_final$LastTransactionDate,format = "%m/%d/%Y")
PropertySales_final$MaturityDate <- as.Date(PropertySales_final$MaturityDate,format = "%m/%d/%Y")

library("lubridate")
PropertySales_final$MaturityDate[is.na(PropertySales_final$MaturityDate)&!is.na(PropertySales_final$LastTransactionDate)]<-PropertySales_final$LastTransactionDate+years(10)

```

##Partioning data into train and test sets 

```{r 5}

set.seed(123)
smp_size<-floor(0.70*nrow(PropertySales_final))
train_ind<-sample(seq_len(nrow(PropertySales_final)),size=smp_size)
train<-PropertySales_final[train_ind,]
#removing TransactionType, LoanMaturityDate and LoanAmount from test set
test<-PropertySales_final[-train_ind,]
str(train)
str(test)

```

##Summary of all missing value information in train and test sets

```{r 6}

#Train Set:
mis_vars_train <- data.frame(colSums(sapply(train, is.na)))
colnames(mis_vars_train) <- 'No_of_NAs'
mis_vars_train
#Test Set:
mis_vars_test <- data.frame(colSums(sapply(test, is.na)))
colnames(mis_vars_test) <- 'No_of_NAs'
mis_vars_test

```

##Replacing NA values in train and test sets using bagImpute method 

```{r 7}

library(caret)
#Train Set
bagImpute_train <-predict(preProcess(train[,which(names(train) %in% 
                  c("NOIAmount", "NumUnits","OrigLoanAmt", "LastTransactionAmount", "LoanCount", "WAInterestRate", "LoanTermMonths", "TotSqrFt", "RentableSqrFt", "AverageRentAmt", "CurOccupancyPct" ))], 
                  method= c("bagImpute")),
                  train[,which(names(train) %in%
                  c("NOIAmount", "NumUnits","OrigLoanAmt", "LastTransactionAmount", "LoanCount", "WAInterestRate", "LoanTermMonths", "TotSqrFt", "RentableSqrFt", "AverageRentAmt", "CurOccupancyPct" ))])

train$NOIAmount <- bagImpute_train$NOIAmount
train$NumUnits <- bagImpute_train$NumUnits
train$OrigLoanAmt <- bagImpute_train$OrigLoanAmt
train$LastTransactionAmount <- bagImpute_train$LastTransactionAmount
train$LoanCount <- bagImpute_train$LoanCount
train$WAInterestRate <- bagImpute_train$WAInterestRate
train$LoanTermMonths <- bagImpute_train$LoanTermMonths
train$TotSqrFt <- bagImpute_train$TotSqrFt
train$RentableSqrFt <- bagImpute_train$RentableSqrFt
train$AverageRentAmt <- bagImpute_train$AverageRentAmt
train$CurOccupancyPct <- bagImpute_train$CurOccupancyPct

#Test Set
bagImpute_test <-predict(preProcess(test[,which(names(test) %in% 
                                                    c("NOIAmount", "NumUnits","OrigLoanAmt", "LastTransactionAmount", "LoanCount", "WAInterestRate", "LoanTermMonths", "TotSqrFt", "RentableSqrFt", "AverageRentAmt", "CurOccupancyPct" ))], 
                                     method= c("bagImpute")),
                          test[,which(names(test) %in%
                                         c("NOIAmount", "NumUnits","OrigLoanAmt", "LastTransactionAmount", "LoanCount", "WAInterestRate", "LoanTermMonths", "TotSqrFt", "RentableSqrFt", "AverageRentAmt", "CurOccupancyPct" ))])

test$NOIAmount <- bagImpute_test$NOIAmount
test$NumUnits <- bagImpute_test$NumUnits
test$OrigLoanAmt <- bagImpute_test$OrigLoanAmt
test$LastTransactionAmount <- bagImpute_test$LastTransactionAmount
test$LoanCount <- bagImpute_test$LoanCount
test$WAInterestRate <- bagImpute_test$WAInterestRate
test$LoanTermMonths <- bagImpute_test$LoanTermMonths
test$TotSqrFt <- bagImpute_test$TotSqrFt
test$RentableSqrFt <- bagImpute_test$RentableSqrFt
test$AverageRentAmt <- bagImpute_test$AverageRentAmt
test$CurOccupancyPct <- bagImpute_test$CurOccupancyPct

#omitting missing date records
train <- na.omit(train)
test<-na.omit(test)

```

##fitting a logistic regression model

```{r 8}

#creating logistic regression model using train data
model1<-glm(TransactionType~.,family=binomial(link="logit"),train)
summary(model1)

#back elimination of predictors having low p-value
train <- train[-12]
test <- test[-12]
model1<-glm(TransactionType~.,family=binomial(link="logit"),train)
summary(model1)
train <- train[, -c(2,7)]
test <- test[, -c(2,7)]
model1<-glm(TransactionType~.,family=binomial(link="logit"),train)
summary(model1)
train <- train[,c(-2,-3)]
test <- test[,c(-2,-3)]
model1<-glm(TransactionType~.,family=binomial(link="logit"),train)
summary(model1)

```

##evaluating the performance on test data: confusionMatrix, ROC and AUC

```{r 9}

p<-predict(model1,test,type="response")
summary(p)

#turn probabilities into classes and look at their frequencies
p_class <- ifelse(p>0.50, "Sale", "Refinance")
table(p_class)
table(test$TransactionType)

#comparing predicted vs actual classes using Confusion Matrix 
confusionMatrix(p_class, test$TransactionType)

#Creating ROC curve
library(caTools)
colAUC(p, test$TransactionType, plotROC = TRUE)

```

##Cross-validation - creating a traincontrol object with 5 folds

```{r 10}

library(caret)
set.seed(42)
myFolds <- createFolds(train$TransactionType, k=5)
myControl <- trainControl(
  summaryFunction = twoClassSummary,
  classProbs = TRUE,
  verboseIter = TRUE,
  savePredictions = TRUE,
  index = myFolds
)

```

##glmnet model - Lasso Regression and Ridge Regression

```{r 11}

set.seed(42)
model_glmnet<-train(TransactionType~.,train, 
                    metric="ROC",
                    method="glmnet",
                    tuneGrid = expand.grid(
                        alpha=0:1,
                        lambda = 0:10/10
                      ),
                      trControl=myControl)

p1<-predict(model_glmnet,test)
confusionMatrix(p1, test$TransactionType)

#plot the results
plot(model_glmnet)

```

##randomforest model        

```{r 12}

library(caret)
set.seed(42)
model_rf<-train(TransactionType~.,train, 
                metric="ROC",
                method="ranger",
                trControl=myControl)

p2<-predict(model_rf,test)
confusionMatrix(p2, test$TransactionType)

#plot the results
plot(model_rf)

```

##comapring glmnet and randomforest models

```{r 13}

model_list <- list(glmnet = model_glmnet, rf=model_rf)
#collect resamples from the CV folds
(resamps <- resamples(model_list))
summary(resamps)
#Box-and-whisker plot
bwplot(resamps, metric="ROC")
#Dot plot
dotplot(resamps, metric="ROC")
#Density plot
densityplot(resamps, metric="ROC")
#Scatter plot
xyplot(resamps, metric="ROC")


```
